---
title: "Ukraine COVID-19 Brief Analysis"
description: |
  Brief analysis about daiuly COVID-19 cases that occur in Ukraine.
author:
  - name: Jimmy Effendy 
    affiliation: Monash University
date: 09-03-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(COVID19)
library(tidyverse)
library(DT)
library(formattable)
library(sparkline)
```

```{css}
thead {
  background-color: darkslategrey;
  color: white;
}
```

```{r data-read}
covid_ukraine <- covid19(country = "Ukraine", verbose = FALSE, raw = FALSE) %>% 
  select(date:deaths) %>% 
  mutate(daily_tests = tests - lag(tests),
         daily_confirmed = confirmed - lag(confirmed),
         daily_recovered = recovered - lag(recovered),
         daily_deaths = deaths - lag(deaths))

covid_ukraine_long <- covid_ukraine %>% 
  select(date, daily_tests, daily_confirmed, daily_recovered, daily_deaths) %>% 
  pivot_longer(cols = daily_tests:daily_deaths,
               names_to = "type",
               values_to = "cases")
```

This blog post focuses in analysing data related to COVID-19 phenomenon that specifically occur in Ukraine.
The main source of data used in this blog originated from `covid19` package.
This data is funded by the Institute for Data Valorization IVADO, Canada.
They collect COVID-19 related data from various parts of the world and merge them to help 
the research community a better understanding of COVID-19 (Institute for Data Valorization IVADO, 2019).

```{r cases}
covid_ukraine %>% 
  select(date, daily_tests, daily_confirmed, daily_recovered, daily_deaths) %>%
  filter(date >= "2020-03-03") %>% 
  datatable(options = list(pageLength = 10),
            colnames = c('ID', 'Date', 'Daily Tests', 'Daily Confirmed Cases', 'Daily Recovered', 'Daily Deaths'),
            caption = 'Table 1: Daily Updates of COVID-19 Statistics in Ukraine')
```

Table 1 shows the number of daily tests, confirmed cases, recovery, and deaths that occurred in Ukraine since 
the day they recorded their first case, which is March 3rd, 2020. As shown in the table, there were no new confirmed 
cases on the first 10 days. However, they started to record COVID-19 related first death on day 11. Starting day 25, 
they started to record new daily cases in the region of 100 cases. By day 46, they recorded a daily jump of 500 cases 
for the first time. The largest number of daily confirmed cases occurred on the 29th of August with 2,579 cases, while 
54 is the largest number of daily deaths which happened on the 3rd of September.

```{r summary-stats}
covid_ukraine_long %>%
  rename("Type" = "type") %>% 
  group_by(Type) %>% 
  summarise(fivenum = list(setNames(as.vector(summary(na.omit(cases))),
                                    c("Min", "Q1", "Median", "Mean", "Q3", "Max"))),
            Distribution = spk_chr(cases, type = "box")) %>% 
  unnest_wider(fivenum) %>% 
  datatable(escape = FALSE,
            caption = "Table 2: Summary Statistics of Daily Cases in Ukraine",
            options = list(paging = FALSE,
                           dom = "t",
                           fnDrawCallback = htmlwidgets::JS(
  '
function(){
  HTMLWidgets.staticRender();
}
'
))) %>% 
  spk_add_deps() %>% 
  formatRound(columns = c("Min", "Q1", "Mean", "Median", "Q3", "Max"), digits = 2)
```

Summary statistics of daily tests, confirmed cases, recovery, and deaths are shown in Table 2.
The table shows that the distributions for these variables are skewed to the right. This means that
the majority of the variables have relatively smaller values, but there are few values are relatively large.
The mean and median of the variables arguably have similar values except for `daily_recovered` where its mean 
are more than twice the value of its median. This means that `daily_recovered` variable has significant influence of
high value outliers. On average (mean), there are 538 people that are tested positive, 253 people recovered, 
and 11 people dead due to the virus in Ukraine. Around 6,646 tests are conducted daily in Ukraine.


```{r correlation}
covid_ukraine %>% 
  select(daily_tests, daily_confirmed, daily_recovered, daily_deaths) %>% 
  mutate(daily_tests = coalesce(daily_tests, 0),
         daily_confirmed = coalesce(daily_confirmed, 0),
         daily_recovered = coalesce(daily_recovered, 0),
         daily_deaths = coalesce(daily_deaths, 0)) %>% 
  transform(daily_tests = as.numeric(daily_tests),
            daily_confirmed = as.numeric(daily_confirmed),
            daily_recovered = as.numeric(daily_recovered),
            daily_deaths = as.numeric(daily_deaths)) %>%
  select(-id) %>%
  cor() %>% 
  as.data.frame() %>% 
  rownames_to_column("Variables") %>% 
  formattable(list(area(col = 2:5) ~ color_tile("#F5B7B1", "#7DCEA0")))
```
